Information Disclosure Concerns in The Age of Wearable Computing

### 2016-01-29 USEC2016-camera-ready/camera-ready.tex
final wordsmith and orphankilling passthrough

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Perceptions of Information Disclosure Risks for Wearable Devices}

@@

\begin{abstract}
\noindent Wearable devices, or ``wearables,'' bring great benefits but also potential information disclosure risks that could expose users' activities without their awareness or consent. We surveyed 1,782 Internet users about various data associated with the capabilities of popular wearable devices on the market to identify which users find most concerning if disclosed. Our study relatively ranks a range of potential data capture scenarios enabled by wearable devices, investigates the impact of the recipient of the data on the perceived risk of data disclosure, and concludes with a survey of users' risks perceptions associated with wearable devices. Users perceive associate risks of owning a wearable with compromised privacy and security, but also consider factors such as safety risk, changes in social behaviors, and how wearable devices are unfashionable. To our knowledge, this is the largest user-based experiment concerning information disclosure regarding wearables. We hope that this work will aid in the design of future user notification, permission management, and access control schemes for wearables.
\end{abstract}

@@

Wearables are a \$700 million, growing industry~\cite{cmo}. With 20\% of the general population owning at least one wearable and 10\% using it daily~\cite{WearableStatNews}, wearables are bringing ubiquitous computing to everyday life. This trend will likely continue, as 52\% of technology consumers are aware of wearables and 33\% are likely to buy one~\cite{NPD}.

@@

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item We give a relative ranking of 72 potential capture scenarios, which were inspired from the capabilities from the most popular wearable devices on the market at the time of the study. 
\item We study 4 possible data recipients to find that the recipient of the data contributes to the magnitude of overall perceived risk, but do not find statistically significant correlated factors of risk between data and recipient. 
\item We sketch a landscape of users' self-reported concerns regarding wearable devices and analyze responses using logistic regression models. 
\end{itemize}

\section{Methodology}
We designed a survey for our IRB-approved study to capture the general public's perception of wearables risks. To determine which of the current data capture scenarios were most concerning to users, we asked them to rate their level of concern for a handful of scenarios from a list of possible scenarios. This was intended to elicit their perception of the severity and impact of the risk. The format of this section was based on Felt {\it et al.}'s study of user perceptions of security and privacy risks with mobile devices~\cite{Felt}. To get a qualitative, unbounded measurement of what people thought the most common risk associated with wearables are, we asked our participants an open-ended question.

To obtain a representative list of scenarios, we examined the sensors, capabilities, permissions, and applications of the most popular wearable devices on the market. At the time of this study (August 2014), the most popular wearable devices included the Fitbit fitness tracker, which continuously monitors heartbeat, steps taken, and sleep patterns; the Pebble smartwatch, which can take pictures, send texts, show notifications from online, and push notifications to services; and Google Glass, which can take pictures, record video, and perform a subset Internet-based tasks such as search, reading emails, etc. These devices' capabilities and requested permissions were the basis for the list of possible data capture scenario questions used in this study, which we feel will be representative of what users are likely to encounter today.

\subsection{Survey Questions}
\noindent We report on participants' responses to 23 questions across 4 survey sections:  \\[-.5cm]

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item 2 reading comprehension questions
\item 6 questions regarding possible wearables scenarios
\item 1 open-ended wearables risk question
\item 14 demographic questions %\\[-.8cm]
\end{itemize}

@@

\noindent {\bfseries Comprehension Questions}
The beginning of the survey introduced participants to the ``Cubetastic3000,'' which was the basis for all questions on wearables risks. Because participants might be biased to specific companies (e.g., visceral reactions to Google Glass based on popular media stories), we framed our scenarios on a fictitious wearable. We highlighted the capabilities of this device and described use cases:

@@

\noindent---measuring heart rate, breathing, and other things to keep track of your fitness level and overall health\\
\noindent---look at what you see to provide information about what's around you\\
\noindent---allow you to take notes just by telling the device what you need to remember\\
\noindent---take videos of you or what you see to share\\
\noindent---automatically take photos or video so that you can replay events that previously happened\\
\noindent---play music that you like for you when it detects that no one is around\\
\noindent---infer information about you so you don't need to log in or search for things\\
\noindent ...and much more!}
\end{quotation}

@@

\noindent {\bfseries Wearable Scenarios}
We presented scenarios involving data captured by the Cubetastic3000 and asked participants to rate how upset they would be if a particular type of data (e.g., how much you exercise) was shared without permission with a particular recipient (e.g., work contacts). The purpose for using this question format was to determine how upset participants would be if data were inappropriately shared, and the extent to which their reactions were based on the data type and recipient. Responses were reported on a 5-point Likert scale (from ``indifferent'' to ``very upset''). Figure \ref{fig:prompt} shows an example.  Specifically, questions were of the form:

\begin{quotation}
\noindent
\textit{``How would you feel if an app on your Cubetastic3000 learned (data) and shared it with (recipient), without asking you first?''}
\end{quotation}

@@

We combined 72 data types, ({\it data}), with 4 data recipients, ({\it recipient}), to form a pool of 288 scenarios (Table \ref{all-vur}). Each participant answered 6 questions that were randomly drawn from this pool, displayed in random order. We clarified for our participants that ``app'' meant the app's server, and that the data was not shared with anyone else.

\subsubsection{Additional Questions}
The exit portion of the survey contained demographic questions (age, gender, and education), and then asked about wearable device ownership so we could control for prior exposure. An open-ended question on what would be the most likely risks associated with wearable devices was included to capture user concerns more broadly. To avoid biasing the open-ended question, we asked before concluding with the 10-question Internet Users' Information Privacy Concerns (IUIPC) index~\cite{malhotra2004internet}, which was included so we could control for participants' general privacy attitudes. However, we do realize that we asked this open-ended question after we had exposed our participants to a variety of wearables scenarios, which may have heightened their awareness of the possible risks. We talk about this more in Section 4.

\subsection{Focus Group}
We conducted a one-hour focus group to validate our design, gauge comprehension, and measure fatigue. The focus group began with participants taking the survey then giving feedback on the format and the content, noting any instructions or questions that were unclear. The focus group concluded with a discussion of possible benefits and risks of wearable devices, in order to brainstorm any additional scenarios to include. Finally, we compensated participants with \$30 in cash. We recruited all of our focus group participants from Craigslist. Of the 13 participants, 54\% were female, and ages ranged from 18 to 64 ($\mu = 36.1$, $\sigma = 15.3$).  Education backgrounds ranged from high school to doctorate degrees, and professions included student, artist, marketer, and court psychologist.

\subsection{Recruitment and Analysis Method}
We recruited 2,250 participants over August 7--13, 2014 via Amazon's Mechanical Turk. We restricted participants to those over 18, living in the United States, and having a successful HIT completion rate of 95\% or above. We compensated each participant with \$1.75 upon successfully completing the survey. Based on incorrect responses to either of the two comprehension questions, we filtered out 366 (16\% of 2,250) participants. We filtered out an additional 99 participants (4\% of 2,250) due to incomplete responses, and three participants for being under 18, leaving us with a total sample size of 1,782. Of these, 57.9\% were male (1,031), 41.0\% were female (731), and 20 participants declined to state their genders. Ages ranged from 18 to 73, with a mean of 32.1 ($\sigma$ = 10.37). Almost half of our participants had completed a college degree or more (49.2\% of 1,782), which includes the 219 (12.3\% of 1,782) who reported graduate degrees. While our sample was younger and more educated than the U.S. population as a whole, we believe it is still consistent with the U.S. Internet-using population.

In performing our analysis in the next section, we chose to focus on the very upset rate (VUR) of each scenario~\cite{Felt}.  The VUR is defined as the percentage of participants who reported a `5' on the Likert scales. 
We use the VURs rather than the average of all Likert scores for the same reasons as Felt {\it et al.}: the VUR does not presume that the ratings, ranging from ``indifferent'' to ``very upset,'' are linearly spaced. Additionally, most people are likely to be upset, at least a little, in all scenarios, because a device is taking action without permission (rating distribution: ``1''= 759, ``2'' = 918, ``3'' = 1,452, ``4''' = 2,421, ``5'' = 8,344). Thus, the distinguishing factor is whether a participant was maximally upset. A limitation of this approach is that it only allows us to make {\it relative} comparisons between scenarios, rather than being able to definitively state how upset people might be if a single scenario were to occur.

\section{Results}
In this section, we present participants' responses to the various data-sharing scenarios, and discuss how and which various factors contributed to their risk perceptions. We had at least 141 responses per data type, 2,779 per recipient, and 35 responses per each unique data type/recipient combination. We conclude the section with participants' self-reported wearables concerns.\\

@@

Participants were most concerned about photos and videos, especially if they contained embarrassing content, nudity, or financial information. As seen in Table \ref{top10-table}, photos and videos accounted for five of the top ten concerns, and are almost unanimously considered to be concerning. Information that could be used to impersonate someone (e.g., usernames/passwords for websites), or photos of someone at home, were also among the most concerning data types.

@@

The statistical significance arises for two distinct reasons. Firstly, sharing data only with an application's server carries less social impact. And for our participants, it may seem that it is shared with fewer people. Additionally, there is a class of data which may be considered odd for a human to know, but completely normal for a wearable device to know (e.g., it's okay if your Fitbit knows when you sleep, but maybe less so for your friends).

We note that this chi-square test violates the assumption of independent observations, since participants responded to multiple scenarios with multiple recipients. But based on the randomization of treatments and large sample size, we do not believe that this significantly impacted our results. Similarly, we are unaware of a more appropriate test, given our data format. Cochran's Q requires binary outcomes (i.e., participants would have had to answer only one question for each data recipient, preventing us from adequately controlling for data type) and a repeated measures ANOVA requires normality (our data was not normally distributed). Nonetheless, we repeated our analysis using only one randomly-selected data point per participant and found that our selected test was robust to this violation. Therefore, we conclude that participants were significantly more concerned about having their data seen by a human versus an application, though differences between specific human groups such as the public, friends, and work contacts were not as significant. Our results motivate mechanisms for data taint tracking and accidental data sharing prevention.

However, we do not claim that there are no distinctions between the friends, public, and work contact recipients. People are more comfortable sharing certain data types with certain human recipients. For instance, participants were significantly more uncomfortable sharing if they were lying, nervous, or stressed to work contacts compared to the rest of the data recipients. Table \ref{all-vur} shows the complete VURs and rankings of all data types by recipient.\\

@@

\begin{quotation}
\noindent
\textit{What do you think are the most likely risks associated with wearable devices?}
\end{quotation}

The participants were presented with a blank box to write in, with no particular suggestion on what to talk about, and with no character limit to their open-ended responses. Table \ref{openresponses} shows common user concerns related to wearable devices. Appendix \ref{sec:coding} details the responses categorized in each coding label. Note that participants were especially concerned with privacy and security, however, we do realize that we asked this open-ended question after we had exposed our participants to a variety of privacy-related scenarios, which may have heightened their awareness of the possible privacy and security risks. \\

@@

Other significant concerns included being unaware of what the device is collecting, doing, or which information it is using (Being Unaware), long-term health effects caused from wearing the device such as cancer from EMF waves (Health), and safety hazards from wearing the device, such as distractions that cause car accidents (Safety).\\

@@

The landscape of users' perceived risks associated with wearable devices is broad, encompassing concepts such as privacy and security, but also non-technical aspects such as health, safety, and change in social behaviors. We hope that this motivates researchers to investigate these aforementioned risks. \\

@@

\noindent {\bfseries Demographics Factors}
We see that privacy is a main concern for wearables users. Additionally, we show that privacy preferences should also be a consideration when configuring a user's device. A participant's self-reported level of privacy concern---as determined by the IUIPC scale~\cite{malhotra2004internet}---is the biggest demographic predictor of VURs. A Spearman correlation yielded a statistically significant effect between average IUIPC scores and average VUR ($\rho=0.446$, $p\textless0.0005$), which suggests responses to questions were mostly based on privacy preferences. Additionally, we observed that age was another significant predictor of VUR ($\rho=0.121$, $p\textless0.0005$), but we suspect that this effect is due to the significant correlation between age and IUIPC scores ($\rho=0.188$, $p\textless0.0005$). Others have observed that older individuals tend to be more privacy protective~\cite{varian2005demographics}.

@@

We created several models using two independent variables as predictors: data and recipient. This resulted in a total of 72 types of data shared with 4 possible recipients. Demographic factors used as covariates are: age, gender, education, wearable device ownership (yes/no), and mean IUIPC score. For each model, we performed Wald's test to examine the model effects attributable to each of these parameters. The only covariates that had an observable effect on our models were participants' gender and IUIPC scores, which also exhibited an interaction effect with each other. Thus, we opted to remove the other covariates from our analysis. Table \ref{regression} shows the various models that we examined and the Quasi-Akaike Information Criterion (QIC), which is a goodness-of-fit metric for model selection that also accounts for complexity (lower relative values indicate better fit). As shown, the type of data being shared (data type) was found to be the strongest predictor of a high VUR.

@@

The data types fell into these six categories:

\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
\item Photo
\item Video
\item Audio
\item Behavioral Information
\item Biometric Information
\item Demographic Information
\end{enumerate}

While the first three categories are self-explanatory, the latter three categories are all based on different user characteristics. We defined {\it behavioral information} as observations about the user's activities; {\it biometric information} as measurements of the user's body; and {\it demographic information} as non-biometric information about the user's traits. \\

The risks for data types fell into these five categories:

\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
\item {\bf Financial:} the loss of money or property.
\item {\bf Image:} the loss of control over one's self-image (e.g., publicizing something embarrassing).
\item {\bf Medical:} the disclosure of medical information.
\item {\bf Physical:} physical harm to the user.
\item {\bf Relationships:} damage to the user's inter-personal relationships.
\end{enumerate}

After independently coding, the researchers met to resolve any disagreements, such that the results reflect unanimity. There was  83\% agreement prior to resolution. Cohen's $\kappa$ was 0.81 for the data categories and 0.75 for the  risk categories, both indicating ``excellent'' agreement~\cite{Fleiss2003}.

With regard to data types, the most concerning type of data is video (78.0\%), which was ranked similarly to photos (76.2\%). Next are audio (66.8\%) and demographic data (65.4\%), followed by behavioral (53.1\%) and biometric (46.3\%) data. We suspect that demographic data was more concerning because it included information such as a Social Security Number, bank account information, and other financial information. We chose to categorize them as such as they are non-biological descriptors of the user. We were very surprised that biometric information was seen as relatively benign compared to the other broad categories of data. One hypothesis is that since most home users do not use biometric authentication, they may have an inaccurate understanding of the types of systems that might be at risk if biometric data is stolen and abused.

With regard to the presented risks, we observed that average VURs were highest for financial information disclosure (82.0\%). Information regarding relationships (69.2\%), physical safety (66.4\%), and self-image (65.8\%) followed. VURs were lowest for medical information disclosure (47.4\%). One reason why medical risks were ranked relatively low is that this category broadly covered scenarios involving data about the user's health, but also included more basic medical information, such as age, gender, and emotional state. As mentioned earlier, participants saw these as publicly observable and therefore unconcerning.

@@

Using these two new variables as additional independent variables (and removing the previous data type variable), we created a second set of models. Because these risk categories and mediums are less likely to change over time, models that take these into account are likely to be more useful and less likely to be overfit. What these models show us is that both risk and medium are relatively strong predictors by themselves, and have an even stronger interaction effect. When the data recipient and covariates are added to the model, the resulting goodness-of-fit is not much worse than that of the model using the actual data type.

@@

\noindent {\bfseries Limitations}
One of the limitations of our experiment is that our participants might not have knowledge or interest in wearables and their capabilities; 83\% of our participants reported that they do not own a wearable. Because of this, our participants may be over or underestimating the risk, stemming from an unawareness of what can be inferred from the data, not having clear relations of new technology with respect to familiar ones, and a higher likelihood of being influenced by reports of recent events.\footnote{At the time of the survey, stories of exploding batteries were in the news~\cite{1_levin_2014}, which were explicitly reported as a concern in our open-ended question.}  For instance, biometrics were generally not a concern for our participants, although there are many security and privacy implications~\cite{prabhakar2003biometric}. Our participants also did not differentiate between the benefits and risks of various new capabilities.

We recruited both wearable users and non-users in order to yield a more representative sample of the general population. We could have easily recruited only wearables owners or people specifically interested in wearables. However, that would have its own biases and limitations. At the time of this writing, about 85\% of the general population do not own wearable devices~\cite{Nilsen,WearableStatNews}, indicating our study is reflective of the current population.

Because of the privacy paradox, participants' stated responses may differ from how they may react to these same scenarios in real life~\cite{norberg2007privacy, jensen2005privacy}. At the same time, our results do reflect actual perceptions of wearable devices and the associated privacy scenarios involving them. This is an unavoidable, yet important distinction to make with of studies of this nature: our primary goal was to examine perceptions and preferences, so that future systems can be designed with these in mind. We do not expect that such systems will satisfy users in all situations, however, we believe that user-centered design will still be a vast improvement over {\it post hoc} approaches (or ignoring user concerns altogether).

@@

\noindent {\bfseries Future Research Directions}
We find that although people have opinions on applications which are familiar, users do not know the risk associated with new data or unfamiliar applications. We hope our work both informs the direction for future research to secure video, audio, and other currently considered sensitive sensor input channels, but also encourages work for contextual and user-input-independent permission models and access control schemes.

Further work can be done to expand various aspects of this study. Investigating more fine-grained data types (e.g., investigating specific instances of location data, versus location data in general) would be a useful endeavor to gain further insight into user perceptions. Adding additional recipients, such as ``advertisers'' or ``acquaintances'' may lead to more nuanced results. Additionally, the open-ended concerns illuminate areas of possible future research, such as the design of a distraction-free interface to prevent safety issues, and how to minimize negative social impact.

@@

One limitation of user perceptions is that people do not always have enough information to make privacy-sensitive decisions. Even if users did have this information, it has been shown that users often trade off long-term privacy for short-term benefits~\cite{acquisti2005privacy}. Furthermore, actual behavior may deviate from stated privacy preferences~\cite{spiekermann2001privacy}. However, understanding user concerns is a necessary first step not only for risk communication, but preventative measures in general against breaches of privacy and security in a new threat landscape. \\

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Information Disclosure Concerns in The Age of Wearable Computing}

@@

\begin{abstract}
\noindent Wearable devices, or ``wearables,'' bring great benefits but also potential information disclosure risks that could expose users' activities without their awareness or consent. We surveyed 1,782 Internet users about various data associated with the capabilities of popular wearable devices on the market to identify which users find most concerning if disclosed. Our study relatively ranks a potential data capture scenarios enabled by wearables and investigates the impact of the recipient of the data on the perceived risk of data disclosure. We conclude with a brief survey of users' perception of  general risks associated with wearable devices, which includes factors such as safety, changes in social behaviors, and impact on fashion. To our knowledge, this is the largest user-based experiment concerning information disclosure regarding wearables. We hope that this work will aid in the design of future user notification, permission management, and access control schemes for wearables.
\end{abstract}

@@

Wearables are a \$700 million, growing industry~\cite{cmo}. With 20\% of the population owning at least one wearable and 10\% using it daily~\cite{WearableStatNews}, ubiquitous computing is becoming a reality. This trend will continue, as 52\% of technology consumers are aware of wearables and 33\% are likely to buy one~\cite{NPD}.

@@

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item We give a relative ranking of 72 potential capture scenarios, which were inspired from the capabilities from the most popular wearable devices on the market at the time of the study. 
\item We study 4 possible data recipients to find that the recipient of the data contributes to the magnitude of overall perceived risk, but do not find statistically significant correlated factors of risk between data and recipient. 
\item We sketch a landscape of users' self-reported general risks regarding wearable devices and analyze responses using logistic regression models. 
\end{itemize}

\section{Methodology}
We designed a survey for our IRB-approved study to capture the general public's perception of information disclosure risks associated with wearables. To determine the most concerning data disclosure scenarios, we asked them to rate their level of concern for a 6 of scenarios from a list of 72 possible scenarios. This was intended to elicit their perception of the severity and impact of the risk. The format of this section was based on Felt {\it et al.}'s study of user perceptions of security and privacy risks with mobile devices~\cite{Felt}. To get a qualitative, unbounded measurement of what people thought the most common risk associated with wearables are, we asked our participants an open-ended question.

To obtain a representative list of scenarios, 4 researchers examined the sensors, capabilities, permissions, and applications of the most popular wearable devices on the market. At the time of this study (August 2014), the most popular wearable devices included the Fitbit fitness tracker, which continuously monitors heartbeat, steps taken, and sleep patterns; the Pebble smartwatch, which can take pictures, send texts, show notifications from online, and push notifications to services; and Google Glass, which can take pictures, record video, and perform a subset Internet-based tasks. These devices' capabilities and requested permissions were the basis for the list of possible data capture scenario questions used in this study, which we feel will be representative of what users are likely to encounter today. For a scenario to be included in this study, the scenario was required to be understandable to a user, possible to happen with today's wearable devices, and capture unique information not in other scenarios.

\subsection{Survey Questions}
\noindent We report on participants' responses to 23 questions across 4 survey sections:  \\

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item 2 reading comprehension questions
\item 6 questions regarding possible wearables scenarios
\item 1 open-ended wearables risk question
\item 14 demographic questions \\
\end{itemize}

@@

\noindent {\bfseries Comprehension Questions}
The survey introduced participants to the ``Cubetastic3000,'' which was the basis for all questions on wearables risks. Because participants might be biased to specific companies (e.g., visceral reactions to Google Glass based on popular media stories), we framed our scenarios on a fictitious wearable. We highlighted the capabilities of this device and described use cases:\\

@@

\noindent---measuring heart rate, breathing, and other things to keep track of your fitness level and overall health\\
\noindent---look at what you see to provide information about what's around you\\
\noindent---allow you to take notes just by telling the device what you need to remember\\
\noindent---take videos of you or what you see to share\\
\noindent---automatically take photos or video so that you can replay events that previously happened\\
\noindent---play music that you like for you when it detects that no one is around\\
\noindent---infer information about you so you don't need to log in or search for things\\
\noindent ...and much more!}\\
\end{quotation}

@@

\noindent {\bfseries Wearable Scenarios}
We presented scenarios involving data captured by the Cubetastic3000 and asked participants to rate how upset they would be if a particular type of data (e.g., how much you exercise) was shared without permission with a particular recipient (e.g., work contacts). The purpose for using this question format was to determine how upset participants would be if data were inappropriately disclosed, and the extent to which their reactions were based on the data type and recipient. Responses were reported on a 5-point Likert scale (from ``indifferent'' to ``very upset''). Figure \ref{fig:prompt} shows an example.  Specifically, questions were of the form: \\

\begin{quotation}
\noindent
\textit{``How would you feel if an app on your Cubetastic3000 learned (data) and shared it with (recipient), without asking you first?''}\\
\end{quotation}

@@

We combined 72 data types ({\it data}) with 4 recipients ({\it recipient}) to form 288 scenarios (Table \ref{all-vur}). Each participant answered 6 randomly drawn questions, displayed in random order. We clarified that ``app'' meant that the data was not shared with anyone else but a server. \\

\noindent {\bfseries Additional Questions}
The exit portion of the survey collected demographics (age, gender, and education) and wearable device ownership so we could control for prior exposure. An open-ended question the most likely risks associated with wearable devices was included to capture user concerns more broadly. To avoid biasing the open-ended question, we asked before concluding with the 10-question Internet Users' Information Privacy Concerns (IUIPC) index~\cite{malhotra2004internet}, which was included so we could control for participants' general privacy attitudes. However, we realize that we asked this open-ended question after exposing participants to a variety of wearables scenarios, which may have heightened their awareness of the possible risks. We talk about this more in Section 4.

\subsection{Focus Group}
We conducted a one-hour focus group to validate our design, gauge comprehension, and measure fatigue. The focus group participants took the survey, gave feedback on the format and the content, and noted any confusions. The focus group concluded with a discussion of possible benefits and risks of wearable devices, in order to brainstorm any additional information disclosure scenarios. The Craigslist-recruited participants received \$30 in cash for their time. Of the 13 participants, 54\% were female, and ages ranged from 18 to 64 ($\mu = 36.1$, $\sigma = 15.3$).  Education backgrounds ranged from high school to doctorate degrees, and professions included student, artist, marketer, and court psychologist.

\subsection{Recruitment and Analysis Method}
We recruited 2,250 participants over August 7--13, 2014 via Amazon's Mechanical Turk. We restricted participants to those over 18, living in the United States, and having a successful HIT completion rate of 95\% or above. We compensated each participant with \$1.75 upon completing the survey. Based on incorrect responses to either of the two comprehension questions, we filtered out 366 (16\% of 2,250) participants. We filtered out an additional 99 participants (4\% of 2,250) due to incomplete responses, and three participants for being under 18, leaving us with a total sample size of 1,782. Of these, 57.9\% were male (1,031), 41.0\% were female (731), and 20 participants declined to state their genders. Ages ranged from 18 to 73, with a mean of 32.1 ($\sigma$ = 10.37). Almost half of our participants had completed a college degree or more (49.2\% of 1,782), which includes the 219 (12.3\% of 1,782) who reported graduate degrees. While our sample was younger and more educated than the U.S. population as a whole, we believe it is still consistent with the U.S. Internet-using population.

In performing our analysis in the next section, we chose to focus on the very upset rate (VUR) of each scenario~\cite{Felt}.  The VUR is defined as the percentage of participants who reported a `5' on the Likert scales. 
We use the VURs rather than the average of all Likert scores for the same reasons as Felt {\it et al.}: the VUR does not presume that the ratings, ranging from ``indifferent'' to ``very upset,'' are linearly spaced. Additionally, most people are likely to be upset, at least a little, in all scenarios, because a device is taking action without permission (rating distribution: ``1''= 759, ``2'' = 918, ``3'' = 1,452, ``4''' = 2,421, ``5'' = 8,344). Thus, the distinguishing factor is whether a participant was maximally upset. A limitation of this approach is that it only allows us to make {\it relative} comparisons between scenarios, rather than being able to definitively state how upset people might be if a single scenario were to occur. This metric is commonly used in marketing research.

\section{Results}
We had at least 141 responses per data type, 2,779 per recipient, and 35 responses per each unique data type/recipient combination. In this section, we present participants' responses to the various data disclosure scenarios and discuss various factors contributed to their risk perceptions. We conclude the section with self-reported general wearables concerns.\\

@@

Participants were most concerned about photos and videos, especially if they contained embarrassing content, personal content, nudity, financial information, or information that can be used to impersonate someone (e.g., usernames/passwords for websites). As seen in Table \ref{top10-table}, photos and videos accounted for five of the top ten concerns, and are almost unanimously considered to be concerning.

@@

The statistical significance arises for two distinct reasons. Firstly, sharing data only with a server carries less social impact. For our participants, it may seem that it is shared with fewer people. Additionally, there is a class of data which may be considered odd for a human to know, but completely normal for a wearable device to know (e.g., it's okay if your Fitbit knows when you sleep, but maybe less so for your friends).

We note that this chi-square test violates the assumption of independent observations, since participants responded to multiple scenarios with multiple recipients. But based on the randomization of treatments and large sample size, we do not believe that this significantly impacted our results. Similarly, we are unaware of a more appropriate test, given our data format. Cochran's Q requires binary outcomes (i.e., participants would have had to answer only one question for each data recipient, preventing us from adequately controlling for data type) and a repeated measures ANOVA requires normality (our data was not normally distributed). Nonetheless, we repeated our analysis using one randomly-selected data point per participant and found that our selected test was robust to this violation. Therefore, we conclude that participants were significantly more concerned about having their data seen by a human than an application, though differences between human groups such as the public, friends, and work contacts were not as significant. Our results motivate mechanisms for data taint tracking and accidental data sharing prevention.

However, we do not claim that there are no distinctions between the friends, public, and work contact recipients. People are more comfortable sharing certain data types with certain human recipients. For instance, participants were significantly more uncomfortable sharing if they were lying, nervous, or stressed to work contacts compared to the rest of the data recipients. We believe a more fine-grained study is required to measure these effects accurately, although we have insight into the existence of these trends. Table \ref{all-vur} shows the complete VURs and rankings of all data types by recipient.\\

@@

\begin{quotation}
\noindent
\textit{What do you think are the most likely risks associated with wearable devices?}\\
\end{quotation}

Participants were asked to write in a blank box, with no particular suggestion on what to talk about or a character limit to their responses. Table \ref{openresponses} shows common user concerns related to wearable devices. Appendix \ref{sec:coding} details which responses were categorized in a coding label. Note that participants were especially concerned with privacy and security. Many answers mention scenarios not included in the study, but prior exposure to our questions may have heightened sensitivity to information disclosure risks. \\

@@

Other common concerns included being unaware of what the device is collecting, doing, or which information it is using (Being Unaware), long-term health effects caused from wearing the device such as cancer from EMF waves (Health), and safety hazards from wearing the device, such as distractions that cause car accidents (Safety).\\

@@

The landscape of users' perceived risks associated with wearables is broad, encompassing concepts outside of privacy and security. We hope that this frames our research to be just one aspect of risks associated with wearables, and motivates researchers to investigate these other risks. \\

@@

\noindent {\bfseries Demographics Factors}
A participant's self-reported level of privacy concern---as determined by the IUIPC scale~\cite{malhotra2004internet}---is the biggest demographic predictor of VURs. A Spearman correlation yielded a statistically significant effect between average IUIPC scores and average VUR ($\rho=0.446$, $p\textless0.0005$), which suggests responses to questions were mostly based on privacy preferences. Age was another significant predictor of VUR ($\rho=0.121$, $p\textless0.0005$), but we suspect that this effect is due to the significant correlation between age and IUIPC scores ($\rho=0.188$, $p\textless0.0005$). Others have observed that older individuals tend to be more privacy protective~\cite{varian2005demographics}.

@@

We created several models using two independent variables as predictors: data and recipient. This resulted in a total of 72 types of data shared with 4 possible recipients. Demographic factors used as covariates are: age, gender, education, wearable device ownership (yes/no), and mean IUIPC score. For each model, we performed Wald's test to examine the model effects attributable to each of these parameters. The covariates that had an observable effect on our models were participants' gender and IUIPC scores, which exhibited an interaction effect with each other. Thus, we opted to remove the other covariates from our analysis. Table \ref{regression} shows the various models that we examined and the Quasi-Akaike Information Criterion (QIC), which is a goodness-of-fit metric for model selection that also accounts for complexity (lower relative values indicate better fit). As shown, the type of data being shared (data type) was found to be the strongest predictor of a high VUR.

@@

The data types fell into these six categories:\\

\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
\item Photo
\item Video
\item Audio
\item Behavioral Information
\item Biometric Information
\item Demographic Information\\
\end{enumerate}

While the first three categories are self-explanatory, the latter three categories are based on user characteristics. We defined {\it behavioral information} as observations about the user's activities; {\it biometric information} as measurements of the user's body; and {\it demographic information} as non-biometric information about the user's traits. \\

The risks for data types fell into these five categories:\\

\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
\item {\bf Financial:} the loss of money or property.
\item {\bf Image:} the loss of control over one's self-image (e.g., publicizing something embarrassing).
\item {\bf Medical:} the disclosure of medical information.
\item {\bf Physical:} physical harm to the user.
\item {\bf Relationships:} damage to the user's inter-personal relationships.\\
\end{enumerate}

After independently coding,  researchers met to resolve any disagreements such that the results reflect unanimity. There was  83\% agreement prior to resolution. Cohen's $\kappa$ was 0.81 for the data categories and 0.75 for the  risk categories, both indicating ``excellent'' agreement~\cite{Fleiss2003}.

With regard to data types, the most concerning type of data is video (78.0\%), which was ranked similarly to photos (76.2\%). Next are audio (66.8\%) and demographic data (65.4\%), followed by behavioral (53.1\%) and biometric (46.3\%) data. We suspect that demographic data was more concerning because it included information such as a Social Security Number, bank account information, and other financial information. We chose to categorize them as such as they are non-biological descriptors of the user. We were surprised that biometric information was seen as  benign. We suspect most users today may have an inaccurate understanding of the risks if biometric data is stolen and abused.

With regard to the presented risks, we observed that average VURs were highest for financial information disclosure (82.0\%). Information regarding relationships (69.2\%), physical safety (66.4\%), and self-image (65.8\%) followed. VURs were lowest for medical information disclosure (47.4\%). One reason why medical risks were ranked relatively low is that this category broadly covered scenarios involving data about the user's health, but also included more basic medical information, such as age, gender, and emotional state.

@@

Using these two new variables as additional independent variables (and removing the previous data type variable), we created a second set of models. Because these risk categories and mediums are less likely to change over time, models that take these into account are likely more useful and less likely to be overfit. What these models show us is that both risk and medium are relatively strong predictors by themselves, and have an even stronger interaction effect. When the data recipient and covariates are added to the model, the resulting goodness-of-fit is not much worse than that of the model using the actual data type.

@@

\noindent {\bfseries Limitations}
One of the limitations of our experiment is that our participants might not have knowledge or interest in wearables and their capabilities; 83\% of our participants reported that they do not own a wearable. Because of this, our participants may be over or underestimating the risk, stemming from an unawareness of what can be inferred from the data, not having clear relations of new technology with respect to familiar ones, and a higher likelihood of being influenced by reports of recent events.\footnote{At the time of the survey, stories of exploding batteries were in the news~\cite{1_levin_2014}, which were explicitly reported as a concern in our open-ended question.}  Biometrics were generally not a concern for our participants, although there are security and privacy implications~\cite{prabhakar2003biometric}. Our participants also did not differentiate between the benefits and risks of various new capabilities.

We recruited both wearable users and non-users to yield a representative sample of the general population. We could have easily recruited only wearables owners or people specifically interested in wearables. However, that would have its own biases and limitations. At the time of this writing, about 85\% of the general population do not own wearable devices~\cite{Nilsen,WearableStatNews}, indicating our study is reflective of the current population.

Because of the privacy paradox, participants' stated responses may differ from how they may react to these same scenarios in real life~\cite{norberg2007privacy, jensen2005privacy}. At the same time, our results do reflect actual perceptions of wearable devices and the associated information disclosure scenarios. This is an unavoidable, yet important distinction to make with of studies of this nature: our primary goal was to examine perceptions and acceptance so that future systems can be designed with these in mind. We do not expect that such systems will satisfy users in all situations, however, we believe that user-centered design will still be a vast improvement over {\it post hoc} approaches (or ignoring user concerns altogether).

@@

\noindent {\bfseries Future Research Directions}
Further work can be done to expand various aspects of this study. A study designed to highlight the differences in information disclosure acceptance, privacy preferences, or technical fluency between wearables owners and non-owners will compliment this work. Investigating more fine-grained data (e.g., investigating specific instances of location data, versus location data in general),  additional recipients (e.g., ``advertisers'' or ``acquaintances'') may lead to more nuanced results.  Additionally, the open-ended concerns illuminate areas of possible future research, such as the design of a distraction-free interface to prevent safety issues, and how to minimize negative social impact.

We find that although people have opinions on applications which are familiar, users are still unfamiliar with particular data types or capabilities. We hope our work both informs the direction for future research to secure video, audio, and other currently considered sensitive sensor input channels, but also encourages work broadly in contextual and user-input-independent permission models and access control schemes.

@@

One limitation of user perceptions is that people do not always have enough information to make privacy-sensitive decisions. Even if users did have this information, it has been shown that users often trade off long-term privacy for short-term benefits~\cite{acquisti2005privacy}. Furthermore, actual behavior may deviate from stated privacy preferences~\cite{spiekermann2001privacy}. However, understanding user concerns is a necessary first step not only for risk communication, but preventative measures against breaches of privacy and security in a new threat landscape. \\
